---
title: "Cheese Data Analysis"
author: "Dakota Chang, Katherine Danielson"
date: 2025-04-17
output:
  github_document:
    toc: true
---

## Cheese Data Intro

**Background Information**

Source: <https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-06-04/readme.md>

*Sourcing and Credibility*

This project uses the dataset from the[ TidyTuesday: Cheese Week (June 4, 2024)](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-06-04/readme.md) initiative. The data is from Cheese.com, a long-standing online encyclopedia for cheeses, and was scraped by data scientist Thomas Mock and Jon Harmon for the purposes of public exploration and education.

This is a popular reference sites and is one of the largest cheese databases in the world, hosting extensive listings and descriptions of cheese worldwide. However, while it hosts a large amount of data, it is not academic nor industry-certified; thus, it introduces uncertainty about standardization and objectivity surrounding entries. Unfortunately, the original site's data-entry methodology is not transparent.

*Potential Errors, Uncertainties and Ambiguities*

With a lack of transparent data entry, this introduces potential errors in data classification. If the website was created through other available data sources or different customer entries, there could be potential errors in classification (e.g., inconsistent naming of regions, vague flavor descriptors) and missing values due to non-uniform entries across cheeses. Additionally, with a lack of transparency, there could be a bias in certain aspects of data entry. For example, it appears in this dataset that the United States of America produces the most kinds of cheese, with France and Italy at different points behind. However, France and Italy are renowned for being some of the top producers of cheeses worldwide with the United States always behind them.

This dataset is a non-random sample, containing \~1,200 cheeses documented on Cheese.com as of 2024. This does not represent the full global population of cheeses but is instead a convenience sample—limited to what has been posted/entered online. Therefore, selection bias is present. Well-known cheeses or those with wider international distribution are likely overrepresented. Regional or artisanal cheeses with less online presence may be underrepresented.

**Question and Exploration**

-   What question did you set out to answer?

-   What data did you find to help answer that question?

## Importing Data and Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
df_cheeses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv')
df_cheeses
```

## EDA and Basic Checks

```{r}
ncol(df_cheeses)
nrow(df_cheeses)
summary(df_cheeses)
```

**Observations:**

-   

The important variables in this dataset are:

| Variable | Class | Description |
|----|----|----|
| cheese | character | Name of the cheese. |
| url | character | Location of the cheese's description at cheese.com |
| milk | character | The type of milk used for the cheese, when known. |
| country | character | The country or countries of origin of the cheese. |
| region | character | The region in which the cheese is produced, either within the country of origin, or as a wider description of multiple countries. |
| family | character | The family to which the cheese belongs, if any. |
| type | character | The broad type or types to describe the cheese. |
| fat_content | character | The fat content of the cheese, as a percent or range of percents. |
| calcium_content | character | The calcium content of the cheese, when known. Values include units. |
| texture | character | The texture of the cheese. |
| rind | character | The type of rind used in producing the cheese. |
| color | character | The color of the cheese. |
| flavor | character | Characteristic(s) of the taste of the cheese. |
| aroma | character | Characteristic(s) of the smell of the cheese. |
| vegetarian | logical | Whether cheese.com considers the cheese to be vegetarian. |
| vegan | logical | Whether cheese.com considers the cheese to be vegan. |
| synonyms | character | Alternative names of the cheese. |
| alt_spellings | character | Alternative spellings of the name of the cheese (likely overlaps with synonyms). |
| producers | character | Known producers of the cheese. |

```{r}
df_cheeses %>%
  group_by(country) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(country, -count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 10 Countries by Cheese Count", x = "Country", y = "Count")
```

**Observations:**

-   

```{r}
df_cheeses %>%
  group_by(milk) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(milk, -count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top Milk Types by Cheese Count", x = "Milk Type", y = "Count")
```

**Observations:**

-   

```{r}
df_cheeses %>%
  mutate(milk = str_split(milk, ", ")) %>%
  unnest(milk) %>%
  group_by(milk) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(milk, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Milk Types by Cheese Count", x = "Milk Type", y = "Count") +
  geom_text(aes(x = milk, y = count,label=count),hjust=-0.1) +
  theme_minimal()

```

**Observations:**

-   

```{r}
df_cheeses %>%
  group_by(type) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(type, -count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 10 Types by Cheese Count", x = "Cheese Type", y = "Count")
```

**Observations:**

-   

```{r}
df_cheeses %>%
  mutate(type = str_split(type, ", ")) %>%
  unnest(type) %>%
  group_by(type) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(type, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Types by Cheese Count", x = "Country", y = "Count") +
  geom_text(aes(x = type, y = count,label=count),hjust=-0.1) +
  theme_minimal()
```

**Observations:**

-   

## Data Analysis

## Conclusion and Remaining Questions
