---
title: "Cheese Data Analysis"
author: "Dakota Chang, Katherine Danielson"
date: 2025-04-25
output:
  github_document:
    toc: true
---

## Cheese Data Intro

#### **Background Information**

Source: <https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-06-04/readme.md>

*Sourcing and Credibility*

This project uses the dataset from the [TidyTuesday: Cheese Week (June 4, 2024)](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-06-04/readme.md) initiative. The data is from Cheese.com, a long-standing online encyclopedia for cheeses, and was scraped by data scientist Thomas Mock and Jon Harmon for the purposes of public exploration and education.

This is a popular reference sites and is one of the largest cheese databases in the world, hosting extensive listings and descriptions of cheese worldwide. However, while it hosts a large amount of data, it is not academic nor industry-certified; thus, it introduces uncertainty about standardization and objectivity surrounding entries. Unfortunately, the original site's data-entry methodology is not transparent.

*Potential Errors, Uncertainties and Ambiguities*

This source introduces a multitude of areas for potential error and uncertainties.

Firstly, as this data was scraped from cheese.com in 2024, there are discontinuities between the cheeses then and now. There is a possibility this led to errors like 5 yak-milk cheeses being present in currently, but only 4 in the 2024 scraping. While this could be due to timing differences, with a lack of tracking on data entry times, this could additionally be due to errors in scraping the web. Consequently, this leads to questions on the legitimacy of the dataset.

Most importantly, there is a lack of transparency surrounding data entry and collection which introduces potential errors in data classification. If the website was created through other available data sources or different customer entries, there could be potential errors in classification (e.g., inconsistent naming of regions, vague flavor descriptors) and missing values due to non-uniform entries across cheeses. Additionally, it appears as though some cheeses are classified as different cheeses when they are only aged differently, while some cheeses are classified the same for different levels of aging. Additionally, with a lack of transparency, there could be a bias in certain aspects of data entry. For example, it appears in this dataset that the United States of America produces the most kinds of cheese, with France and Italy at different points behind. However, France and Italy are renowned for being some of the top producers of cheeses worldwide with the United States always behind them. As this is a website that sells cheeses and/or offers locations to sell cheeses as well as boasts its unique cheese collection, collection bias is introduced. It is likely that cheeses that are overly common are highly present as a site of cheeses would be "incomplete" without them. Further, as they boast their unique and broad collection, they also likely highlight very artisan and varied cheeses.

This dataset is a non-random sample, containing \~1,200 cheeses documented on Cheese.com as of 2024. This does not represent the full global population of cheeses but is instead a convenience sample—limited to what has been posted/entered online. Therefore, selection bias is present. Well-known cheeses or those with wider international distribution are likely overrepresented. Regional or artisanal cheeses with less online presence may be underrepresented. Additionally, due to the fact that this sampling is not random, it would be unhelpful to perform confidence and prediction intervals as data would be skewed. Instead, source criticism (as seen above) is better to clarify the uncertainty seen in this dataset.

#### **Population**

-   We will be looking at the non-random sample (1187 cheese) from the cheese.com TidyTuesday dataset.

#### **Quantity of interest**

-   Our main quantity of interest is the country. This recognizes where a cheese is being produced.

#### **Question / Hypothesis**

-   How are the top five cheese-producing countries and cheese qualities, namely milk type, color and cheese type, related?

    -   Null hypothesis: There is no statistical difference between the top five cheese-producing countries based on certain cheese qualities.

        -   There is no statistical difference between the top five cheese-producing countries based on type of cheese.

        -   There is no statistical difference between the top five cheese-producing countries based on the milk used in cheese.

        -   There is no statistical difference between the top five cheese-producing countries based on the color of cheese.

## Importing Data and Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(broom)
```

```{r}
df_cheeses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-06-04/cheeses.csv')
df_cheeses
```

## EDA and Basic Checks

```{r}
ncol(df_cheeses)
nrow(df_cheeses)
summary(df_cheeses)

df_cheeses %>% 
  distinct(cheese)

sapply(df_cheeses, function(x) sum(is.na(x)))  # Count missing values per column

```

**Observations:**

-   This dataset has 19 columns and 1187 rows.

    -   As each cheese only has one set of information associated with it, that means there are 1187 distinct cheeses in the dataset. Thus, each cheese is only found in one county/region.

-   The majority of data in the dataset is in character form.

    -   Things like `type`, `texture` and `flavor` are all characters separated by commas. Thus, data analysis can both be performed with the data as it is and data where the lists are wrangled apart.

-   There are a lot of missing datapoints in a variety of columns.

    -   The only two columns that have no missing information are `cheese` and its associated `url`.

    -   Comparatively, `alt_spellings` has the highest number at 1078. This indicates that 1078 of the cheeses do no have any other spellings, which should not impact data analysis.

    -   However, there are some more valuable variables that are missing a high quantity of their data such as `fat_content` and `region` and if a cheese is `vegetarian`.

-   The majority of cheese recorded are not `vegan` while roughly 50% of the cheeses that report if they are `vegetarian` show that they are.

The important variables in this dataset are:

| Variable | Class | Description |
|------------------------|------------------------|------------------------|
| cheese | character | Name of the cheese. |
| url | character | Location of the cheese's description at cheese.com |
| milk | character | The type of milk used for the cheese, when known. |
| country | character | The country or countries of origin of the cheese. |
| region | character | The region in which the cheese is produced, either within the country of origin, or as a wider description of multiple countries. |
| family | character | The family to which the cheese belongs, if any. |
| type | character | The broad type or types to describe the cheese. |
| fat_content | character | The fat content of the cheese, as a percent or range of percents. |
| calcium_content | character | The calcium content of the cheese, when known. Values include units. |
| texture | character | The texture of the cheese. |
| rind | character | The type of rind used in producing the cheese. |
| color | character | The color of the cheese. |
| flavor | character | Characteristic(s) of the taste of the cheese. |
| aroma | character | Characteristic(s) of the smell of the cheese. |
| vegetarian | logical | Whether cheese.com considers the cheese to be vegetarian. |
| vegan | logical | Whether cheese.com considers the cheese to be vegan. |
| synonyms | character | Alternative names of the cheese. |
| alt_spellings | character | Alternative spellings of the name of the cheese (likely overlaps with synonyms). |
| producers | character | Known producers of the cheese. |

### Exploration of Country

```{r}
df_cheeses %>%
  group_by(country) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(country, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Top 10 Countries by Cheese Count", x = "Country", y = "Count")+
  theme_minimal()
```

**Observations:**

-   In this database, the United States has a significantly higher number of cheese types represented compared to other countries. The U.S. lists over 300 types, while the second-highest, France, has just over 150—meaning the U.S. has nearly twice as many. This discrepancy could be due to a few factors. The most likely explanations are: (1) the U.S. genuinely produces a wider variety of cheeses, or (2) the source of the data—a cheese e-commerce platform—may be based in the U.S. or have more partnerships with U.S.-based producers.

### Exploration of Milk Types

```{r}
df_cheeses %>%
  mutate(milk = str_split(milk, ", ")) %>%
  unnest(milk) %>%
  group_by(milk) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(milk, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Milk Types by Cheese Count", x = "Milk Type", y = "Count") +
  geom_text(aes(x = milk, y = count,label=count),hjust=-0.1) +
  theme_minimal()
```

**Observations:**

-   It seems that the overwhelming majority of cheese is made from cow's milk, with goat's milk being the second most common. This is not surprising, as cow's milk is the most widely produced milk in the world. Sheep's milk is the third most common, but it is still significantly less common than cow's and goat's milk. This is likely due to the fact that sheep's milk is less widely available and more expensive than cow's and goat's milk. We also see a few interesting milks, for example, there is one cheese made of moose milk, one of donkey milk, one of camel milk, and four of yak milks. We also see that plant-based milks are listed, but not with any specifications of which kind. It is important to note that this graph removes the detail of cheese made with a few different milks.

```{r}
df_cheeses %>%
  filter(str_detect(milk, ", ")) %>%  # Only rows with multiple milk types
  group_by(milk) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(milk, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Cheeses Made with Mixed Milk Types", x = "Milk Combination", y = "Count") +
  geom_text(aes(label = count), hjust = -0.1) +
  theme_minimal()
```

**Observations:**

-   The most common combination of milk types is cow and goat, which is not surprising given that these are the two most common types of milk used in cheese production. We also see that there are two cheeses that are made from four types of milk (cow, goat, sheep, water buffalo).

### Exploration of Cheese Types

```{r}
df_cheeses %>%
  group_by(type) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(type, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Top 10 Types (non-separated) by Cheese Count", x = "Cheese Type", y = "Count")+
  theme_minimal()

```

**Observations:**

-   The most common cheese type is semi-hard and artisan. The top four all include artisan as a quality, which is not a regulated term, and might be used as a marketing tactic when it comes to types.

```{r}
df_cheeses %>%
  mutate(type = str_split(type, ", ")) %>%
  unnest(type) %>%
  group_by(type) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(type, -count), y = count)) +
  geom_bar(stat = "identity", fill = '#eecc11') +
  coord_flip() +
  labs(title = "Types by Cheese Count", x = "Cheese Type", y = "Count") +
  geom_text(aes(x = type, y = count,label=count),hjust=-0.1) +
  theme_minimal()
```

**Observations:**

-   We see that artisan is the most popular `type` of cheese. This is then followed by soft and semi-soft, suggesting that they are additionally popular kinds of cheese on cheese.com. When looking at the graph "Top 10 Types (non-separated) by Cheese Count" we can see that the top four cheese `types` all include some form of artisan cheese type; thus, making the 635 counts of artisan cheese highly reasonable. Additionally, that graph shows a wide combination of different types of soft and semi-soft cheeses, indicating it is a popular blended type or characterization of cheese.

```{r}
# stacked bar chart cheese vs milk types ordered by number of cheese types
df_cheeses %>%
  mutate(
    milk_list = str_split(milk, ", ")
  ) %>%
  unnest(milk_list) %>%
  mutate(
    type_list = str_split(type, ", ")
  ) %>%
  unnest(type_list) %>%
  group_by(milk_list, type_list) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>% 
  group_by(milk_list) %>% 
  mutate(
    n_total = sum(count),
    frac = count / n_total,
    n_categories = n()
  ) %>% 
  ggplot(aes(x = reorder(milk_list, n_categories), fill = type_list)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(
    title = "Milk Type vs Cheese Type Composition",
    x = "Milk Type", y = "Percentage", fill = "Cheese Type"
  ) +
  theme_minimal()

```

**Observations:**

-   Less popular/common milk types, such as donkey, moose, camel, and plant-based, have many fewer types of cheeses that they are made into. It is possible that certain types of milk only have properties conducive to certain types of cheese. Comparatively, milks that are more common, like cow, goat and sheep milk are able to be used in a lot more types of cheese.

## Analysis

This section includes analysis for the null hypothesis and its three sub-hypotheses.

```{r}
# stacked bar chart of milk type in top 5 country

df_cheeses %>%
  mutate(
    milk_list = str_split(milk, ", ")
  ) %>%
  unnest(milk_list) %>%
  filter(country %in% c("United States", "France", "Italy", "Canada", "Australia")) %>%
  ggplot(aes(x = country, fill = milk_list)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(
    title = "Top 5 Countries by Cheese Milk Type Composition",
    x = "Country", y = "Percentage", fill = "Milk Type"
  ) +
  theme_minimal()

#Chi-Squared test overall
# Build contingency table from raw counts
milk_country_table <- df_cheeses %>%
  mutate(milk_list = str_split(milk, ", ")) %>%
  unnest(milk_list) %>%
  filter(country %in% c("United States", "France", "Italy", "Canada", "Australia")) %>%
  count(country, milk_list) %>%
  pivot_wider(names_from = milk_list, values_from = n, values_fill = 0) %>%
  column_to_rownames("country") %>%
  as.matrix()

# Perform Chi-squared test
milk_chisq_result <- chisq.test(milk_country_table)

# Output result
milk_chisq_result

#Chi-Squared between each type of milke
top_countries <- df_cheeses %>%
  count(country, sort = TRUE) %>%
  slice_head(n = 5) %>%
  pull(country)

# Prepare the dataset with individual milk types
df_milk_split <- df_cheeses %>%
  filter(country %in% top_countries) %>%
  mutate(milk_list = str_split(milk, ", ")) %>%
  unnest(milk_list)

# Get unique milk types
milk_types <- df_milk_split %>%
  distinct(milk_list) %>%
  pull()

# Perform a chi-squared test for each milk type across countries
milk_chi_tests <- map_df(milk_types, function(milk_type) {
  df_test <- df_milk_split %>%
    mutate(has_milk = milk_list == milk_type) %>%
    count(country, has_milk) %>%
    pivot_wider(names_from = has_milk, values_from = n, values_fill = 0)

  # Skip if not enough variation
  if (ncol(df_test) < 3 || sum(df_test[[2]]) == 0 || sum(df_test[[3]]) == 0) {
    return(NULL)
  }

  tbl <- as.matrix(df_test[,-1])
  rownames(tbl) <- df_test$country

  chisq_result <- chisq.test(tbl)
  tidy(chisq_result) %>%
    mutate(milk_type = milk_type)
})

# View results
milk_chi_tests %>%
  select(milk_type, statistic, p.value) %>%
  arrange(p.value)
```

**Observations:**

*"Top 5 Countries by Cheese Milk Type Composition"*

-   This stacked bar chart shows that certain countries bias developing cheeses with certain types of milk. For example, Italy has a lot more cheeses made with sheep milk compared to all other countries. However, there can be trends seen for the most common kinds of milk across all countries. That is, in all top five countries, cow milk is the most common type of milk to use in cheeses. With the exception of Italy (which uses a lot more sheep and water buffalo milk than any other country), goat milk and then sheep milk are the next two most popular kinds of milk in cheeses. It is important to note that the actual percentage allocations between all the countries shift quiet variably.

-   The United States, Italy, Canada and Australia all appear to use at least some percent of buffalo, cow, goat, sheep and water buffalo milk and have at least some milks where the cheese was not reported (NA). Comparatively, cheeses from France appear to have all milk documented for all of their cheeses (no NA values) and only use three milks: goat sheep and cow.

*Chi-Squared Tests*

-   When performing a single chi-squared test across countries with their milk counts, the chi-squared across all cheeses and the top five countries shows a very low p value and a very high X-squared value, indicating significant difference.

-   The second chi-squared test looks at each individual cheese's variation. All of them still have a p-value beneath 0.01 and have a very high X-squared value. The most prominent statistical difference is seen through the proportionate use of sheep milk. When referencing the *Top 5 Countries by Cheese Milk Type Composition* we see that this is echoed with the high level of variation seen between the size of blue bars across countries. Overarchingly, it is reasonable to say there is a statistical difference between the top five countries producing cheese and their milk proportion use.

```{r}
# stacked bar chart of cheese type in top 5 country
df_cheeses %>%
  mutate(
    type_list = str_split(type, ", ")
  ) %>%
  unnest(type_list) %>%
  filter(country %in% c(
    "United States",
    "France",
    "Italy",
    "Canada",
    "Australia"
  )) %>%
  group_by(country, type_list) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(country, -count), fill = type_list)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(
    title = "Top 5 Countries by Cheese Type Composition",
    x = "Country", y = "Percentage", fill = "Type"
  ) +
  theme_minimal()

#Chi-Squared test overall
# Build contingency table from raw counts
type_country_table <- df_cheeses %>%
  mutate(type_list = str_split(type, ", ")) %>%
  unnest(type_list) %>%
  filter(country %in% c("United States", "France", "Italy", "Canada", "Australia")) %>%
  count(country, type_list) %>%
  pivot_wider(names_from = type_list, values_from = n, values_fill = 0) %>%
  column_to_rownames("country") %>%
  as.matrix()

# Perform Chi-squared test
type_chisq_result <- chisq.test(type_country_table)

# Output result
type_chisq_result

#Chi-Squared between each chees type across countries
# Prepare the dataset with individual milk types
df_type_split <- df_cheeses %>%
  filter(country %in% top_countries) %>%
  mutate(type_list = str_split(type, ", ")) %>%
  unnest(type_list)

# Get unique milk types
type_types <- df_type_split %>%
  distinct(type_list) %>%
  pull()

# Perform a chi-squared test for each milk type across countries
type_chi_tests <- map_df(type_types, function(type_type) {
  df_type_test <- df_type_split %>%
    mutate(has_type = type_list == type_type) %>%
    count(country, has_type) %>%
    pivot_wider(names_from = has_type, values_from = n, values_fill = 0)

  # Skip if not enough variation
  if (ncol(df_type_test) < 3 || sum(df_type_test[[2]]) == 0 || sum(df_type_test[[3]]) == 0) {
    return(NULL)
  }

  tbl <- as.matrix(df_type_test[,-1])
  rownames(tbl) <- df_type_test$country

  chisq_result <- chisq.test(tbl)
  tidy(chisq_result) %>%
    mutate(type_type = type_type)
})

# View results
type_chi_tests %>%
  select(type_type, statistic, p.value) %>%
  arrange(p.value)
```

**Observations:**

*Top 5 Countries by Cheese Type Composition*

-   When looking at the visualization "Top 5 Countries by Cheese Type Composition," there is a very large span of cheese types and the presence of each of them.

-   All countries have a wide distribution of cheese types, with very few cheese types not represented many times. One of the few cheeses that follow this exception is whey cheese which is only seen in France and Australia. Some countries have greater level of cheese type reportings than other countries. Australia, France and United States all have NA reportings that are roughly equal to one type of cheese. However, Italy and Canada both have little to no NA reportings. This lack of NAs allows there to be much greater quantities in different cheese types. For example, Italy has a much larger amount of soft-ripened cheese than other countries. However, as there are so many different cheese types and such a large quantity of cheeses reported in the dataset, this graph is very hard to determine. Thus, a chi-squared analysis was performed to try to better understand the variation between cheese types across the top five different countries.

*Chi-Squared Table*

-   The chi-squared table shows that there is a statistical difference between different countries and certain cheese types. The overall chi-squared test shows a high value in the X-squared value and a very low p-value, indicating high statistical difference. To break this analysis down further, we performed a chi-squared for each individual cheese type across the different countries. With a 0.05 threshold, every single cheese still had a high X-squared value and low p-value and was able to remain statistically different. Out of all of the cheese types, soft is by far the most statistically different and soft-ripened and fresh soft come soon after. This indicates that there is large variance in the soft cheese category between countries which makes sense as it is one of the most popular/produces types of cheese.

```{r}
top_countries <- df_cheeses %>%
  count(country, sort = TRUE) %>%
  slice_head(n = 5) %>%
  pull(country)

# Then, filter the data to only include those countries and group by cheese_color
df_cheeses %>%
  filter(country %in% top_countries) %>%
  group_by(country, color) %>%
  summarise(count = n(), .groups = "drop") %>%
  ggplot(aes(x = reorder(country, -count), y = count, fill = color)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Country",
    y = "Number of Cheeses",
    title = "Top 5 Cheese-Producing Countries by Cheese Color",
    fill = "Cheese Color"
  ) +
  theme_minimal()

#Chi-Squared test overall
# Build contingency table from raw counts
color_country_table <- df_cheeses %>%
  mutate(color_list = str_split(color, ", ")) %>%
  unnest(color_list) %>%
  filter(country %in% c("United States", "France", "Italy", "Canada", "Australia")) %>%
  count(country, color_list) %>%
  pivot_wider(names_from = color_list, values_from = n, values_fill = 0) %>%
  column_to_rownames("country") %>%
  as.matrix()

# Perform Chi-squared test and output result
color_chisq_result <- chisq.test(color_country_table)
color_chisq_result

#Run a Chi Squared test for each color across the different top five countries
# Top 5 countries with most cheeses
top_countries <- df_cheeses %>%
  count(country, sort = TRUE) %>%
  slice_head(n = 5) %>%
  pull(country)

# Filter cheese data to top countries only
df_filtered <- df_cheeses %>%
  filter(country %in% top_countries)

# List of cheese colors
cheese_colors <- df_filtered %>%
  distinct(color) %>%
  pull()

# Run chi-squared test per color, safely
color_tests <- map_df(cheese_colors, function(clr) {
  df_color <- df_filtered %>%
  filter(color == clr)

  # Contingency table of country counts
  tbl <- table(df_color$country)

  # Skip if less than 2 non-zero entries
  if (sum(tbl > 0) < 2) return(NULL)

  # Run test and return tidy result
  chisq_result <- chisq.test(tbl)
  tidy(chisq_result) %>%
  mutate(color = clr)
})

# View results
color_tests %>%
  select(color, statistic, p.value)
```

**Observations:**

*"Top 5 Cheese-Producing Countries by Cheese Color"*

-   When looking at the visualization "Top 5 Cheese-Producing Countries by Cheese Color" there is a very large span of cheese colors and the presence of each of them.

-   Three colors as well as NAs appear to be prominent across all of the top five countries: pale yellow, white, and ivory. While these three cheese are all mostly prominent (but to a lesser extent in Australia), there is still a large variation between all of the countries. This could be for a variety of different reasons. This could be influenced by the milk and type palette of the cheese as well as the overall production value. The latter is largely true as we see generally an overarching decreasing trend as we go from the U.S. to France to Italy to Canada to Australia with the cheese colors. However, there are several exceptions. One of these exceptions is the minimal/diminished presence of white cheese in Canada, where the coloring is more distributed, while Australia has mainly white cheeses. Another exception is pale yellow cheese and to an extent ivory cheese. This cheese is seen in abundance in the United States, but much less so in other countries.

-   It is important to note that NA reportings for cheese colors appear to be especially high in places like France. This could lead to a skewing of data and make it hard to interpret if the values are actually statistically different.

*Chi-Squared Table*

-   The chi-squared table shows that there is a statistical difference between different countries and certain cheese colors. The three cheese colors exhibiting the highest difference are ivory, white and pale yellow (as seen in the prior observations). These cheese colors have both a very low p-value and a very high chi-squared statistic value, indicating a large difference between countries. When setting a 0.05 significance value 6/13 cheeses fall above this list, indicating that due to their higher p-value and low statistic value, there is not a statistical difference between cheese color and country for them. This may be due to a smaller sample size/presence of these cheeses in general, however. Overall, it can be said that there is a statistical difference between certain cheese colors across the top five countries that produce cheese.

## Conclusion and Remaining Questions

***Conclusions***

Our guiding question for this analysis was: "How are the top five cheese-producing countries and cheese qualities, namely milk type, color and cheese type, related?" In turn, this led to the null hypotheses of:

-   There is no statistical difference between the top five cheese-producing countries based on certain cheese qualities.
    -   There is no statistical difference between the top five cheese-producing countries based on type of cheese.
    -   There is no statistical difference between the top five cheese-producing countries based on the milk used in cheese.
    -   There is no statistical difference between the top five cheese-producing countries based on the color of cheese.

From our key analyses about we reject the null hypothesis "There is no statistical difference between the top five cheese-producing countries based on certain cheese qualities."

Instead, as proven through three different chi-squared analyses and data visualization for each of our three more specific null hypotheses, there is statistical difference between the top five cheese-producing countries, the types of cheese they make, the color the cheeses and milk used in them.

When looking at the graph *"Top 5 Countries by Cheese Milk Type Composition,"* and its consequent chi-squared analysis, we reject the null hypothesis "There is no statistical difference between the top five cheese-producing countries based on the milk used in cheese." The overall chi-squared analysis shows an X-squared value of 193.21 and a p-value of 2.2\*10\^-6. These values indicate statistical difference. While there are similar trends of what milk is deemed "most popular" or "most used," the actual proportion of its use between each country varies widely. Further, After the top 1-2 milk types, the countries vary widely on choice.

When looking at the graph "*Top 5 Countries by Cheese Type Composition"* and its consequent chi-squared analysis, we reject the null hypothesis "There is no statistical difference between the top five cheese-producing countries based on type of cheese. Instead, its overall chi-squared analysis indicates an X-squared value of 313.94 and a p-value of 2.2\*10\^-6, indicating statistical variance. When further breaking the analysis down to look at each cheese type across each country, all p-values remained below 0.05. Most countries have a very wide variety of cheese types that they use. More importantly, the most statistically significant groups tend to have a common genre of type. That is, soft is by far the most statistically different and soft-ripened and fresh soft come soon after. All three cheeses belong to the "soft" grouping, and if these three were lumped together, or cheese were grouped otherwise (ex. hard, firm, ripened, etc) there could be differences in the results.

Lastly, when looking at the graph *"Top 5 Cheese-Producing Countries by Cheese Color,"* and its consequent chi-squared analysis, we reject the null hypothesis "There is no statistical difference between the top five cheese-producing countries based on the color of cheese." Instead, when looking at its chi-squared analysis, we find that 7/13 cheese colors have statistical difference across the countries, as they have p-values below 0.05 and a high X-squared value. The overall chi-squared value has a p-value of 1.918\*10\^-14 and an X-squared value of 190.15. While color is statistically different amongst countries, the chi-squared analysis is likely highly influenced by the pale yellow, ivory, white and yellow p-values. There is large variation between main colors, but it diminishes as color popularity decreases.

However, despite rejecting all three null hypotheses, the results may not be fully trusted. The source for which this data comes from calls into several questions and possible sources of error that could influence these findings. One such source of error/influence comes from the selection of cheese data seen on the cheese.com website. As cheese.com boasts a wide range of cheeses, it is plasusible that they may pull a wide variety of cheeses. In doing so, they could curate a cheese database with a wider variety of milk types, cheese types and colors than typical. This in turn would make data seem a lot more statistically different than it actually is. Additionally, as cheese.com includes no indication of popularity of production and only shows that it is produced in a certain country, this could influence the statistical difference as well. For example, the number of types of cheese with straw coloring is very low, but if it is a highly popular cheese, it could be the most produced cheese color in a country. Due to the lack of production quantification, there is no way to tell this. Other areas of error include the fact that each cheese is produced in only one country, leading to limitations in variance, as well as a lack of reporting and NA values. Some times, the NA values account for the same quantity as a factor in a variable does. If these NA values had reportings, it is plausible that differences could shift. Due to these reasons, while we still reject the null hypotheses, we recommend proceeding with caution before designing models or making estimates based on this non-random sample.

***Remaining Questions***

After performing this analysis, several questions still remain.

As this analysis only focused on the top five cheese-producing countries to make the analysis comprehensible, it would be interesting to *analyze other countries further*. This could lead to several questions:

-   "Do these statistical differences continue when analyzing only smaller cheese-producing countries?"

-   "How does variation in cheese type, milk and color change when countries produce less cheese?"

Other questions could look at *more aspects of cheese or different regions*:

-   "How are country and cheese qualities, namely milk type, flavor and cheese type, related?"

-   "How are regions in a specific country and cheese qualities, namely milk type, flavor and cheese type, related?"

-   "How are the country and cheese qualities, namely cheese flavor, family and texture, related?"

As the *method of data collection and categorization* is unclear it led to several other questions:

-   "How was the color and flavor of cheese determined? Is there a standardized way that these were categorized?"

-   "Why can a cheese only be produced in one country?"

-   "Where did cheese.com collect this data from and how?"

This project led to many interesting insights. With a dataset with as many variables as this cheese dataset, there are many more questions that can be asked to further explore possible relations and differences with cheeses and their production.
